---
title: "DS311 - R Lab Assignment"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: united
    highlight: tango
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
num_variables <- ncol(mtcars)
num_observations <- nrow(mtcars)

# Answer:
print("There are total of _____ variables and _____ observations in this data set.")
```



b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!


# Answer:
print("There are 11 discrete variables and 32 continuous variables in this data set.")
```



c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!
m <- mean(mtcars$mpg)
v <- var(mtcars$mpg)
s <- sd(mtcars$mpg)



print(paste("The average of Mile Per Gallon from this data set is ", m , " with variance ",  v, " and standard deviation",s  , "."))
```



d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!
install.packages("dplyr")
library(dplyr)

# Table 1: Average mpg for each cylinder class
avg_mpg_by_cyl <- mtcars %>%
  group_by(cyl) %>%
  summarise(Average_MPG = mean(mpg))
# Table 1: Print
print(avg_mpg_by_cyl)

# Table 2: Standard deviation of mpg for each gear class
sd_mpg_by_gear <- mtcars %>%
  group_by(gear) %>%
  summarise(Standard_Deviation_MPG = sd(mpg))
# Table 2: Print
print(sd_mpg_by_gear)



```



e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
crosstab <- table(mtcars$cyl, mtcars$gear)
print(crosstab)

max_count <- max(crosstab)
most_common_combination <- which(crosstab == max_count, arr.ind = TRUE)

num_cylinders <- most_common_combination[1, 1]
num_gears <- most_common_combination[1, 2]
total_cars <- max_count

print("The most common car type in this data set is car with 3 cylinders and 1 gears. There are total of 12 cars belong to this specification in the data set.")
```


*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!
install.packages("ggplot2")
library(ggplot2)

ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  labs(title = "Comparison of Plant Weights by Group",
       x = "Group",
       y = "Weight (grams)") +
  theme_minimal()

# My findings
# Control is pretty good, trt1 is worse than controll so dont do that, trt2 is better than controll.



```

Result:

=> Report a paragraph to summarize your findings from the plot!



b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}
data(mtcars)
hist(mtcars$mpg, breaks = 20, 
     main = "Histogram of Miles Per Gallon (MPG)", 
     xlab = "Miles Per Gallon (MPG)", 
     ylab = "Frequency", 
     col = "lightblue", 
     border = "darkblue")

most_observed_mpg <- as.numeric(names(which.max(table(mtcars$mpg))))


print("Most of the cars in this data set are in the class of about 16 mile per gallon.")
```



c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)

# Enter your code here!
pairs(USArrests, 
      main = "Pairs Plot of US Arrests Data",
      pch = 19, 
      col = "lightblue")

plot(USArrests$Murder, USArrests$Assault,
     main = "Scatter Plot of Murder vs. Assault",
     xlab = "Murder Rate",
     ylab = "Assault Rate",
     pch = 19, 
     col = "darkblue")

```

Result: As the Assault Rate goes up also the murder rate increases.

=> Report a paragraph to summarize your findings from the plot!



***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='data/housing.csv', mode='wb')
housingData <- read.csv('data/housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)

# Enter your code here!
summary_stats <- housingData %>%
  summarise(
    Mean_Value = mean(Market.Value.per.SqFt),
    Median_Value = median(Market.Value.per.SqFt),
    Max_Value = max(Market.Value.per.SqFt),
    Min_Value = min(Market.Value.per.SqFt)
  )

avg_value_neighborhood <- housingData %>%
  group_by(Neighborhood) %>%
  summarise(Average_Value = mean(Market.Value.per.SqFt))

avg_value_borough <- housingData %>%
  group_by(Boro) %>%
  summarise(Average_Value = mean(Market.Value.per.SqFt))

print(summary_stats)
print(avg_value_neighborhood)
print(avg_value_borough)


```

b. Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!
ggplot(housingData, aes(x = Year.Built, y = Market.Value.per.SqFt)) +
  geom_point(alpha = 0.5) +
  labs(title = "Market Value per SqFt vs Year Built",
       x = "Year Built",
       y = "Market Value per SqFt") +
  theme_minimal()

ggplot(housingData, aes(x = Boro, y = Market.Value.per.SqFt)) +
  geom_boxplot() +
  labs(title = "Market Value per SqFt by Borough",
       x = "Borough",
       y = "Market Value per SqFt") +
  theme_minimal()

```

c. Write a summary about your findings from this exercise.

=> It shows that Manhattan has the highest Market Value per SqFt!

Have a nice day!  ( ͡° ͜ʖ ͡°)


